Modeling the impacts of park access on health outcomes: a utility-based accessibility approach

### 2020-09-22 manuscript/submission/manuscript.tex
Submit first revisions

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

@@

\runninghead{Macfarlane \emph{et al}.}

@@

This study uses data available to the public from a variety of federal
and state data agencies\footnote{The datasets, as well as the analysis
  code, are available on GitHub at
  \url{https://github.com/gregmacfarlane/parks_access}.}. The Centers
for Disease Control and Prevention makes small-area estimates on key
health indicators available through its 500 Cities data program
\citep{CDC5002016}. The indicators are multilevel aggregations and
imputations of BRFSS responses \citep{Wang2018, Wang2017}, and have been
recently used to study the tract-level link between gentrification and
urban health \citep{Gibbons2018}. We use two indicators as our dependent
variables in this study: the share of adults in a Census tract who are
obese, and the share of adults who participate in no leisure-time
physical activity. To improve clarity in our interpretation, we use the
complement of the second variable --- the share of tract adults who
participate in \emph{some} physical activity --- even if the amount may
not be sufficient to affect overall health. Both indicators are obtained
for the year 2016.

@@

To determine which spatial effects specification was appropriate, we
estimated SLX and SDEM models of physical activity participation rates
as a function of tract covariates and accessibility. These two models
are presented in Table \ref{tab:slx-sdem}. With only a few exceptions,
the SLX and SDEM coefficient estimates for corresponding covariates lie
within the 95\% confidence intervals of the other model. A Hausman-style
test \citep{Pace2008} on the coefficients assumes a null hypothesis that
the coefficients are equivalent; this test produces \(p\)-value of only
\(1.307e-23\). Though this statistic is well beneath a common hypothesis
test threshold of \(\alpha = 0.05\), we elect to retain the null hypothesis
that the coefficients are equivalent, as the differences are at no point
practically different from each other. The consequences of rejecting the
null hypothesis in this case would be to suggest a model with spatial
dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we
adopt the SDEM specification moving forward as the spatial lag of the
errors is significant and the overal log likelihood of the SDEM is
higher than the SLX specification.

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and
accessibility coefficients for three models using different
specifications of the utility equation, as well as the ten-minute walk
buffer calculated by ParkScore. Note that the model called ``Size and
Distance'' is precisely the model labeled ``SDEM'' in Table
\ref{tab:slx-sdem}. The controlling covariates are suppressed for
clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in the Appendix. Each of the
three utility-based specifications is significant at the
\(\alpha = 0.05\) level or lower, and is of the hypothesized
directionality. That is, residents living in tracts with increased
utility-based accessiblity to green spaces show a significantly higher
physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible
tract (\(A_i = -3\)) to the most-accessible tract (\(A_i = 3\)) is
expected to raise the physical activity rate by 0.9 percentage points.
With the understanding of the full 95\% confidence interval, we could
suggest that the true effect of excellent versus poor park access is
anywhere between 0.06 and 1.8 percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the \emph{obesity} rate in a tract, and physical activity becomes an
independent covariate alongside the controlling variables and the
accessibility metrics. Table \ref{tab:obesity-models} presents estimates
of the effect of all three utility-based accessibility measures as well
as the 10-minute walk buffer. As before, the estimates of the other
variables are available in the Appendix. The correlation between
physical activity rates and obesity is clear; for every percentage point
increase in physical activity rate, the expected obesity rate drops by
\(0.43\) percentage points in all four accessiblity specifications. The
indirect effect of physical activity rates on obesity in neighboring
zones is inconsequential, supporting our methodological assertion that
public health variables are not spatially dependent, though the factors
leading to them may be.

@@

\bibliographystyle{sageh.bst}
\bibliography{library.bib}

\hypertarget{appendix}{%
\subsection{Appendix}\label{appendix}}

In this appendix we present the complete estimation results for the
models relating different definitions of access to physical activity (in
Table \ref{tab:pa-fullmodels}) and to obesity (in Table
\ref{tab:ob-fullmodels}). In each case we also present a base model with
no accessibility statistics for comparison.

>>>>

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
\renewcommand{\appendixname}{Supplementary Material}

@@

\runninghead{XXXX \emph{et al}.}

@@

This study uses data available to the public from a variety of federal
and state data agencies\footnote{The datasets, as well as the analysis
  code, are available on GitHub at
  \url{https://github.com/XXXX}.}. The Centers
for Disease Control and Prevention makes small-area estimates on key
health indicators available through its 500 Cities data program
\citep{CDC5002016}. The indicators are multilevel aggregations and
imputations of BRFSS responses \citep{Wang2018, Wang2017}, and have been
recently used to study the tract-level link between gentrification and
urban health \citep{Gibbons2018}. We use two indicators as our dependent
variables in this study: the share of adults in a Census tract who are
obese, and the share of adults who participate in no leisure-time
physical activity. To improve clarity in our interpretation, we use the
complement of the second variable --- the share of tract adults who
participate in \emph{some} physical activity --- even if the amount may
not be sufficient to affect overall health. Both indicators are obtained
for the year 2016.

@@

To determine which spatial effects specification was appropriate, we
estimated SLX and SDEM models of physical activity participation rates
as a function of tract covariates and accessibility. These two models
are presented in Table \ref{tab:slx-sdem}. With only a few exceptions,
the SLX and SDEM coefficient estimates for corresponding covariates lie
within the 95\% confidence intervals of the other model. A Hausman-style
test \citep{Pace2008} on the coefficients assumes a null hypothesis that
the coefficients are equivalent; this test produces \(p\)-value of only
\(1.307e-23\). Though this statistic is well beneath a common hypothesis
test threshold of \(\alpha = 0.05\), we elect to retain the null hypothesis
that the coefficients are equivalent, as the differences are at no point
practically different from each other. The consequences of rejecting the
null hypothesis in this case would be to suggest a model with spatial
dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we
adopt the SDEM specification moving forward as the spatial lag of the
errors is significant and the overall log likelihood of the SDEM is
higher than the SLX specification.

@@

Table \ref{tab:pa-models} presents the estimated spatial lag and
accessibility coefficients for three models using different
specifications of the utility equation, as well as the ten-minute walk
buffer calculated by ParkScore. Note that the model called ``Size and
Distance'' is precisely the model labeled ``SDEM'' in Table
\ref{tab:slx-sdem}. The controlling covariates are suppressed for
clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in Table \ref{tab:pa-fullmodels}. Each of the
three utility-based specifications is significant at the
\(\alpha = 0.05\) level or lower, and is of the hypothesized
directionality. That is, residents living in tracts with increased
utility-based accessibility to green spaces show a significantly higher
physical activity participation rate. Taking the rough mean of the 2
estimated utility-based coefficients, moving from the least-accessible
tract (\(A_i = -3\)) to the most-accessible tract (\(A_i = 3\)) is
expected to raise the physical activity rate by 0.9 percentage points.
With the understanding of the full 95\% confidence interval, we could
suggest that the true effect of excellent versus poor park access is
anywhere between 0.06 and 1.8 percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the \emph{obesity} rate in a tract, and physical activity becomes an
independent covariate alongside the controlling variables and the
accessibility metrics. Table \ref{tab:obesity-models} presents estimates
of the effect of all three utility-based accessibility measures as well
as the 10-minute walk buffer. As before, the estimates of the other
variables are available in Table \ref{tab:ob-fullmodel} in the Supplementary Material. The correlation between
physical activity rates and obesity is clear; for every percentage point
increase in physical activity rate, the expected obesity rate drops by
\(0.43\) percentage points in all four accessibility specifications. The
indirect effect of physical activity rates on obesity in neighboring
zones is inconsequential, supporting our methodological assertion that
public health variables are not spatially dependent, though the factors
leading to them may be.

@@

\bibliographystyle{sageh.bst}
\bibliography{manuscript.bbl}

\hypertarget{appendix}{%
\subsection{Supplementary Material}\label{appendix}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}

In this supplementary material we present the complete estimation results for the
models relating different definitions of access to physical activity (in
Table \ref{tab:pa-fullmodels}) and to obesity (in Table
\ref{tab:ob-fullmodels}). In each case we also present a base model with
no accessibility statistics for comparison.
